Action Groupoids and Fractional Types

### 2016-07-05 pibackground.tex
secs 1 2 3 4

As an example, let $\mathbb{3}$ abbreviate the type
$(\ot \oplus \ot) \oplus \ot$. There are three values of type
$\mathbb{3}$ which are $\inl{\inl{\unitv}}$, $\inl{\inr{\unitv}}$, and
$\inr{\unitv}$. There are six combinators of type
$\mathbb{3} \iso \mathbb{3}$, each representing a different
permutation of three elements, which could be written as follows:
\[\begin{array}{rcl}
a_1 &=& \idiso \\
a_2 &=& \swapp \oplus \idiso \\
a_3 &=& \assocrp \odot (\idiso \oplus \swapp) \odot \assoclp \\
a_4 &=& a_2 \odot a_3 \\
a_5 &=& a_3 \odot a_2 \\
a_6 &=& a_4 \odot a_2
\end{array}\]
Tracing the evaluation of $a_2$ on each of the possible inputs yields:
\[\begin{array}{rcl}
\evalone{(\swapp\oplus\idiso)}{\inl{\inl{\unitv}}} &=& \inl{\evalone{\swapp}{\inl{\unitv}}} \\
&=& \inl{\inr{\unitv}} \\
\\
\evalone{(\swapp\oplus\idiso)}{\inl{\inr{\unitv}}} &=& \inl{\evalone{\swapp}{\inr{\unitv}}} \\
&=& \inl{\inl{\unitv}} \\
\\
\evalone{(\swapp\oplus\idiso)}{\inr{\unitv}} &=& \inr{\evalone{\idiso}{\unitv}} \\
&=& \inr{\unitv}
\end{array}\]
Thus the effect of combinator $a_2$ is to swap the values
$\inl{\inl{\unitv}}$ and $\inl{\inr{\unitv}}$ leaving the value
$\inr{\unitv}$ intact. Iterating $a_2$ again is therefore equivalent
to the identity permutation, which can be verified using the 
2-combinators:
\[\begin{array}{rcl}
a_2 \odot a_2 &=& (\swapp \oplus \idiso) \odot (\swapp \oplus \idiso) \\
&\isotwo& (\swapp \odot \swapp) \oplus (\idiso \odot \idiso) \\
&\isotwo& \idiso \oplus \idiso \\
&=& \idiso
\end{array}\]

>>>>

As an example, let $\mathbb{3}$ abbreviate the type
$(\ot \oplus \ot) \oplus \ot$. There are three values of type
$\mathbb{3}$ which are $\inl{\inl{\unitv}}$, $\inl{\inr{\unitv}}$, and
$\inr{\unitv}$. There are, up to equivalence, six combinators of type
$\mathbb{3} \iso \mathbb{3}$, each representing a different
permutation of three elements, which could be written as follows:
\[\begin{array}{rcl}
a_1 &=& \idiso \\
a_2 &=& \swapp \oplus \idiso \\
a_3 &=& \assocrp \odot (\idiso \oplus \swapp) \odot \assoclp \\
a_4 &=& a_2 \odot a_3 \\
a_5 &=& a_3 \odot a_2 \\
a_6 &=& a_4 \odot a_2
\end{array}\]
Tracing the evaluation of $a_2$ on each of the possible inputs yields:
\[\begin{array}{rcl}
\evalone{(\swapp\oplus\idiso)}{\inl{\inl{\unitv}}} &=& \inl{\evalone{\swapp}{\inl{\unitv}}} \\
&=& \inl{\inr{\unitv}} \\
\\
\evalone{(\swapp\oplus\idiso)}{\inl{\inr{\unitv}}} &=& \inl{\evalone{\swapp}{\inr{\unitv}}} \\
&=& \inl{\inl{\unitv}} \\
\\
\evalone{(\swapp\oplus\idiso)}{\inr{\unitv}} &=& \inr{\evalone{\idiso}{\unitv}} \\
&=& \inr{\unitv}
\end{array}\]
Thus the effect of combinator $a_2$ is to swap the values
$\inl{\inl{\unitv}}$ and $\inl{\inr{\unitv}}$ leaving the value
$\inr{\unitv}$ intact. Iterating $a_2$ again is therefore equivalent
to the identity permutation, which can be verified using the 
2-combinators:
\[\begin{array}{rcl}
a_2 \odot a_2 &=& (\swapp \oplus \idiso) \odot (\swapp \oplus \idiso) \\
&\isotwo& (\swapp \odot \swapp) \oplus (\idiso \odot \idiso) \\
&\isotwo& \idiso \oplus \idiso \\
&=& \idiso
\end{array}\]
